---
title: "link_street_trees_hex_holc"
author: "Dexter H. Locke, PhD"
date: "`r format(Sys.time())`"
output: html_document
editor_options: 
  chunk_output_type: console
---

This *.Rmd does some of the initial data wrangling associated with a paper and analyses led by Karin Burhardt that analyzes street trees in Baltimore, MD and redlning.

[KARIN CITATION HERE]

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 0 set up
```{r}

# Load libraries
packs <-c(  'tidyverse' # cuz
          , 'tidylog'   # prints out what was done in dplyr and tidyr; VERBOSE
          , 'magrittr'  # for all of the the pipes
          , 'sf'        # simple features spatial data suppport and analyses
          , 'janitor'   # pipe-friendly cross tabulations
          ,'tidycensus' # access to Census data in a tidy way
          , 'mapview'   # quick webmaps
          # , 'parzer'    # parses messy coordinates
          , 'tmap'      # thematic mapper
          , 'tictoc'    # timing things.
          )

# # check for all of the libraries
if (length(setdiff(packs, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packs, rownames(installed.packages())))
}

# load the libraries
lapply(packs, library, character.only = TRUE) # raster masks many tidyverse function, be explicit!

# HOLC custom colors
holc_col <- c('#92BC6B', '#92C7C9', '#E7DC6B', '#E47D67')
holc_col2<- c('#92BC6B', '#92C7C9', '#E7DC6B', '#E47D67', '#808080')


d_path <- file.path(getwd(), 'input_data')
```



# 1 read in data
## A HOLC polys 
```{r}
(rl <- st_read('../BHall/HOLC_X_Sites/HOLC_Baltimore/shp/HOLC_Baltimore.shp') %>% 
    select(holc_id, holc_grade) %>% 
    st_transform(crs = 4326) %>% 
    st_cast('POLYGON') %>% 
    mutate(neigh_area = st_area(.)) %>% 
    rowid_to_column(var = 'poly_id'))

```



## B Baltimore street trees
```{r}

st <- st_read('../../evo/AFRI_local/input_data/bc_forestry_trees_12232020/bc_forestry_trees_12232020.shp', quiet = F) %>% 
  st_transform(., crs = st_crs(rl)) 


# exploratory cross-tabulations
st %>% st_drop_geometry() %>% tabyl(SPACE_TYPE)
st %>% st_drop_geometry() %>% tabyl(LOC_TYPE)
st %>% st_drop_geometry() %>% tabyl(Side)
st %>% st_drop_geometry() %>% tabyl(SPACE_TYPE, COMMON)
  

```



## C get Baltimore City boundary
```{r}

census_api_key('db9b3481879b9e79eb8c86608656c3c8a8640bbb', install = TRUE, overwrite = TRUE)

# set mapping options
# this makes loading the data faster on the second/subsquent version.
options(tigris_use_cache = TRUE)


baci <- get_acs(state = 'MD',                 # Maryland
                    county = c('Baltimore City'), # The CITY is its own county
                    geography = 'county',   
                    year = 2017,                 
                    survey = 'acs5',             
                    moe_level = 95, 
                    variables = c(tot_pop = 'B02001_001'), 
                    output = 'wide',
                    geometry = TRUE) %>% 
  st_transform(crs = st_crs(rl)) %>% # reproject to match HOLC polygons
  rename(county = 'NAME')


mapview(baci)

# # write it out for cartography
# baci %>% 
#   st_write(., paste0(getwd(), '/output_data/city_limits_', Sys.Date(),'.shp'))
```



## D hexbins!
### i make them
```{r}
# (rl %>% st_drop_geometry() %>% select(area) %>% min() -> min_holc_area)
# (rl %>% st_bbox() %>% st_as_sfc() %>% st_as_sf() %>% 
#   mutate(area = st_area(.),
#          length = st_length(.)) -> bbox_dims)

# sizes based on trial and error
# cell_size <- 0.001 
# cell_size <- 0.0011
# cell_size <- 0.0012
# cell_size <- 0.0013
# cell_size <- 0.0014
#cell_size <- 0.00145
#cell_size <- 0.0015
cell_size <- 0.002
# cell_size <- 0.0005 # makes many smaller polygons


rl %>%
  st_make_grid(., cellsize = cell_size, square = FALSE,
               #flat_topped = TRUE
               ) %>% 
  st_as_sf() -> hex

# # do you want to check>
# mapview(rl, zcol = 'holc_grade', col.region = holc_col) + 
#   mapview(hex, alpha.regions = 0)
```



### ii filter to ensure they are within The City
```{r}

hex %<>% st_join(.
                , baci
                , left = FALSE
                , join = st_within
              ) 

hex %>% mapview()


```

#### a how much area of BACI is holc (hex)?
```{r}

rl %>%
  st_intersection(baci) %>%
  mutate(clipped_area = as.numeric(st_area(.))) -> holc_in_baci_clipped

mapview(holc_in_baci_clipped) + mapview(baci)

baci %>%mutate(clipped_area = st_area(.)) |> pull(clipped_area) -> approx_baci_area


round(100*(sum(holc_in_baci_clipped$clipped_area) / as.numeric(approx_baci_area)), 2)


holc_in_baci_clipped |> 
  group_by(holc_grade) |> 
  summarise(area_x_grade = sum(clipped_area)
            , pct_area = 100*(area_x_grade / approx_baci_area))

```



### iii join hexbins to holc polygons
```{r}

hex %>%
  st_join(., rl,
          left = TRUE,
          join = st_within,
          # largest = TRUE
          ) %>% 
  filter(!is.na(holc_id)) %>% 
  rowid_to_column(var = 'hex_id') %>% 
  mutate(hex_area = round(st_area(.)),
         holc_id.hex_id = paste(holc_id, hex_id, sep = '.')) %>% 
  # goal is to max trees per hex (therefore larger), but too big and the hexes per neigh shrinks
  tidylog::filter(holc_id != 'D' & holc_id != 'C' & holc_id != 'B14' & holc_id != 'B22') -> hex_holc

# save it out
hex %>%
  st_join(., rl,
          left = TRUE,
          join = st_within,
          # largest = TRUE
          ) %>%
  filter(!is.na(holc_id)) %>% 
  st_write(., paste0(getwd(), '/output_data/hex_', Sys.Date(),'.shp'))




# rl %>% 
#   #filter(holc_id == 'C4' | holc_id == 'C5') %>% 
#   mapview(., zcol = 'holc_grade', col.region = holc_col) +
#   mapview(hex_holc, alpha.region = 0, #lwd =2, 
#         color = 'black')

rl %>% filter(holc_id == 'D6' | holc_id == 'B22') %>% mapview(., zcol = 'holc_grade', col.region = holc_col)

# rl %>% filter(holc_id == 'B14' | holc_id == 'B22') %>% mapview(., zcol = 'holc_grade', col.region = holc_col)

# B16
# rl %>% filter(holc_id == 'B6' | holc_id == 'B7' | holc_id == 'B16') %>% 
#   mapview(., zcol = 'holc_grade', col.region = holc_col) + mapview(hex_holc)

# # exploratory
# rl %>% filter(holc_id == 'A2') %>% 
#   mapview(., zcol = 'holc_grade', col.region = holc_col) + mapview(hex_holc)

# hex_holc %>% 
#   #filter(holc_id == 'C') %>% 
#   filter(!is.na(holc_id)) %>% 
#   mapview(., alpha.region = 0)
```


### iv join / filter street trees 
```{r}

st %>% 
  st_join(., hex_holc, left = TRUE) %>% 
  # st_join(., hmt, left = TRUE) %>% # old, from an exploratory analysis
  filter(!is.na(holc_id.hex_id)) -> st_hex

# a double check
length(table(st_hex$holc_id))

# # write it out for carto
# hex %>%
#   st_write(., paste0(getwd(), '/output_data/hex_', Sys.Date(),'.shp'))



# # write it out
# st_hex %>%
#   st_drop_geometry() %>%
#   write_csv(., paste0(getwd(), '/output_data/street_trees_Baltimore_w_HOLC_grades_HEX_', Sys.Date(),'.csv'))


rl %>% st_cast('POLYGON') -> rl_single

rl_single %>% st_drop_geometry() %>% tabyl(holc_id) %>% 
  arrange(desc(n)) %>% head(15)
```



# 2 maps
## A redlining polygons
```{r}
# TODO scale bar/text
# TODO leader
#rl %>% st_cast('POLYGON') %>% st_centroid() -> holc_pts # %>% mapview() 
# tm_shape(holc_pts) + tm_dots(title = 'holc_id')

rl %>% mutate(' ' = 'Outside Baltimore City') %>% tm_shape(.) + tm_polygons(' ', palette = 'gray') +
st_intersection(rl, baci) %>% tm_shape(.) +
  tm_polygons('holc_grade', palette = holc_col, title = 'HOLC Grade') + # 1937
  tm_text('holc_id', size = .5, shadow = TRUE) +
#rl %>% mutate(' ' = 'Outside Baltimore City') %>% tm_shape(.) + tm_polygons(' ', palette = 'gray') +
tm_shape(baci) + tm_borders(col = 'black') + 
  tm_layout('I.', #legend.stack = 'horizontal',
            legend.title.size=.9,
            legend.text.size = .75
            # legend.bg.alpha = .5,
            # legend.bg.color = 'white'
            ) -> p1

p1 # look

tmap_save(p1, w = 1050*4, h = 450*4 , 'figures/p1_holc_city.png')
```


## B Hexes
```{r}
ex_poly <- rl %>% filter(holc_id == 'D1' | holc_id == 'C1' | holc_id == 'C3')
tm_shape(ex_poly) + tm_polygons() + # sets extent
  # tm_polygons('holc_grade', palette = holc_col[3:4], title = 'HOLC Grade') + 
  # tm_text('holc_id', size = .5, shadow = TRUE)
tm_shape(rl) + tm_polygons('holc_grade', palette = holc_col, legend.show = FALSE) + 
  tm_text('holc_id', size = 3, col = 'black') +
tm_shape(hex_holc) + tm_borders(col = 'black') + tm_layout('II.', title.bg.color = 'white') -> p2

p2 # look

tmap_save(p2, w = 1050*4, h = 450*4 , 'figures/p2_select_neighs.png')
```


## C street trees
```{r}
ex_hex_id_list <- c('C1.1538', 'C1.1547', 'C1.1553', 'C1.1546', 'C1.1537', 'C1.1542', 'C1.1533')

mid <- 'C1.1542'
ex_hex <- hex_holc %>% filter(holc_id.hex_id %in% ex_hex_id_list)

ext <- ex_hex %>% filter(holc_id.hex_id == mid) %>% st_bbox() %>% st_as_sfc() %>% st_buffer(.0009)
#tm_basemap('Esri.WorldImagery') +
ex_hex %>% tm_shape(., bbox = ext) + tm_polygons(alpha = .5) +
ex_hex %>% filter(holc_id.hex_id == mid ) %>% tm_shape(.) + tm_borders(col = 'red') +
tm_shape(ex_hex) + tm_borders(col = 'black') +
st_hex %>% filter(holc_id == 'C1') %>% tm_shape(.) + tm_dots(title = 'street trees') +
  tm_layout('III', title.bg.color = 'white') -> p3


p3 # look

tmap_save(p3, w = 1050*4, h = 450*4 , 'figures/p3_select_hexs.png')
```

# End


# save
```{r eval=FALSE, include=FALSE}
system.time(save.image(file = paste0('saved_sessions/link_street_trees_hex_holc_',
                                     gsub('[[:punct:]]', '-', Sys.time()),
                                     '.RData')))
```


# CITE your sources!!!
```{r eval=FALSE, include=FALSE}
lapply(packs, citation); rm(packs)
sessionInfo()
```

Last Knit on `r format(Sys.time())`